---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# Create time series of years to estimate population size 
time_yr <- seq(from = 0, to = 8, by = 1)

# Create an empty matrix to store pop sizes 
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# Populating first row with year 0 population sizes [row, column]
proj_year[1, ] <- fish_year0

# Calculate projected population based on previous time and fills in matrix with pop estimates
for (i in 2:length(time_yr)) { #starts at 2 b/c row 1 is already filled in
 #can define i and run the inside of the for loop to test what its computing
 # i <-2
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}

# The rest is data wrangling and data visualization 
# Changes column names to life stages
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

# Convert to a data frame 
proj_df <- data.frame(time_yr, proj_year) %>% 
  # Uses pivot longer to convert format from wide to long
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

#Uses data from the df to plot each life stage population at each year
ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 
